Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2020-12-15 manuscript/draft_rsfmri.Rmd
minor changes in manuscript

### Aggregated FC values
To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and time points, adjusted for age and sex. We used the modified SwE option and used global signal regressed data as input. Network masks were formed from all voxels with a clusterwise $p_{FWE} < 0.05$ and we extracted the average connectivity from these masks for all data points.

@@

<!-- MODEL SETUP --> As specified in the preregistration, we tested the interaction of time and group: $FC = group + time + group*time$. Unlike prespecified, we did not include the model without covariates (as this is deprecated \@ref{nuisance_paper}). Instead, we calculated a model adjusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2). We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
These were examined each for correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $16$ a priori planned analyses.
Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$. As a result a total of $32$ different analyses were performed.
Computation with the grey matter mask were exploratory and deviate from the preregistration.
The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.

@@

FDFC=calc_qc_fc(final)
write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)
```

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.

@@

### DMN for two timepoints
```{r calc DMN two tp, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
```
### DMN for two timepoints, adjusted for age, sex, mean FD
```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
```
### DMN for three timepoints, adjusted for age, sex, mean FD
```{r calc DMN three tp, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group ++ age + sex + meanFD (1|subj))
```

### reward
```{r calc rew two tp, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(rew_FC ~ timepoint*group + (1|subj))
R0 = lme4::lmer(rew_FC ~ timepoint + group + (1|subj))
```
### reward for two timepoints, adjusted for age, sex, mean FD
```{r calc rew two tp adjusted, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
```
### reward for three timepoints, adjusted for age, sex, mean FD
```{r calc rew three tp, eval = FALSE, echo = FALSE}
R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
```

>>>>

### Aggregated FC values
To extract aggregated FC values, we first calculated the mean DMN and reward networks over all participants and timepoints, adjusted for age and sex. We used global signal regressed data as input and clusterwise bootstrapping with N=999. Network masks were formed from all voxels with a clusterwise p_FWE < 0.05 and we extracted the average connectivity from these masks for all data points.

@@

<!-- MODEL SETUP --> As specified in the preregistration, we tested the interaction of time and group: $FC = group + time + group*time$. Unlike prespecified, we did not include the model without covariates (as this is deprecated \@ref{nuisance_paper}). Instead, we calculated a model adjusting for age and sex (Model 1) and another model additionally adjusting for mean FD (Model 2). We did not adjust for baseline BMI as described in the preregistration, but explored the relationship with BMI in a separate analysis.
The model was examined for the correlation maps of two different seed regions (NAcc / PCC), and for the purpose of comparison for correlation maps of two different preprocessing pipelines (only AROMA+CC / with GSR)), with two different covariate sets (with/ without mean FD), resulting in $8$ a priori planned analyses.

Initially, analyses were run with a brain mask but after post hoc considerations detailed below, models were re-run with a grey matter mask which reduced the number of tests (voxels of interest) from $52378$ to $29451$. As a result a total of $32$ different analyses were performed.
Computation with the grey matter mask were exploratory and deviate from the preregistration.
The second model on the influence of between-subject differences in average BMI and longitudinal within-subject change in BMI was specified as follows: $FC = between-subject BMI + within-subject BMI$.

@@

FDFC=calc_qc_fc(final)
write.csv(FDFC,"./report/FDFCcorrelations.csv", row.names = FALSE)

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction will be published on NeuroVault.

@@

### DMN for two timepoints

```{r calc DMN two tp, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC ~ timepoint*group + (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group + (1|subj))
```
### DMN for two timepoints, adjusted for age, sex, mean FD
```{r calc DMN two tp adjusted, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD (1|subj))
```
### DMN for three timepoints, adjusted for age, sex, mean FD
```{r calc DMN three tp, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group ++ age + sex + meanFD (1|subj))
```

### reward
```{r calc rew two tp, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(rew_FC ~ timepoint*group + (1|subj))
R0 = lme4::lmer(rew_FC ~ timepoint + group + (1|subj))
```
### reward for two timepoints, adjusted for age, sex, mean FD
```{r calc rew two tp adjusted, eval = FALSE, echo = FALSE}
R1 = lme4::lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
R0 = lme4::lmer(DMN_FC ~ timepoint + group + age + sex + meanFD(1|subj))
```
### reward for three timepoints, adjusted for age, sex, mean FD
```{r calc rew three tp, eval = FALSE, echo = FALSE}
R1 = lmer(DMN_FC ~ timepoint*group + age + sex + meanFD (1|subj))
```
